---
title: "Main Figure: Admixture mapping using local ancestry counts in RFMix and FLARE3 at the chromosome with most significant admixture mapping associations for the four metabolites: 3-aminoisobutyrate at chromosome 5 (A), N-acetylarginine at chromosome 5 (B), PE 16:0/20:4 at chromosome 15(C), and PC 16:0/20:4 at chromosome 11 (D)."
output: html_document
date: "2024-07-23"
---

```{r setup, include=FALSE, fig.width=16}
knitr::opts_chunk$set(echo = TRUE)
```

*Load packages*
```{r}
library(ggplot2)
library(gridExtra)
library(tidyverse)
library(reshape2)
library(lattice)
library(grid)
library(gridExtra)
library(gridtext)
```

```{r}
# Load the liftOver files for interval info in the RFMix
intervals_liftover_nodup <- read.table("/Volumes/Sofer Lab/HCHS_SOL/Projects/2023_local_ancestry_comparison_sol/Data/intervals_37-38.bed")

colnames(intervals_liftover_nodup) <- c("chr", "pos_start", "pos_end", "snpid") 
intervals_liftover_nodup <- intervals_liftover_nodup |> filter(chr != "chrX") # 14753 mapped blocks
```

**Define a function that overlaps the RFMix and FLARE3 admixture mapping result from specified ancestry, batch, and chromosome**
```{r}
plot_overlay_pval <- function(metab, batch, ancestry, chr){
  # Load the RFMix admixmap results 
  res_RFMix <- readRDS(paste0("./admix_map_all/RFMix_res_", ancestry,
  "_", metab, "_b", batch, ".rds"))
  res_RFMix_plot <- cbind(intervals_liftover_nodup$chr, 
                          intervals_liftover_nodup$pos_start,
                          intervals_liftover_nodup$pos_end,
                          res_RFMix[order(as.numeric(rownames(res_RFMix))),])
  colnames(res_RFMix_plot)[1:3] <- c("chr", "pos.start", "pos.end")
  res_RFMix_plot_long <- pivot_longer(res_RFMix_plot, 
                                           cols = c("pos.start", "pos.end"), 
                                           names_to = "poscat", values_to = "pos")
  # Load the FLAREs3 admixmap results
  load(paste0("./admix_map_all/admixmap_", metab, "_all_FLARE3/admix", ancestry,
              "_b", batch, "_filtered_chr", chr, ".Rdata"))
  
  # Start plotting
  # determine colors based on the ancestry given in the argument
  if(ancestry == "afr"){
    colors <- c("FLARE3" = "#AADCA9", "RFMix" = "#24500F")
    legend_name <- "African Ancestry"
  }else{
    colors <- c("FLARE3" = "#ffb38a", "RFMix" = "#854900")
    legend_name <- "Amerindian Ancestry"
  }
  chrnum <- paste0("chr", chr)
  overlay_plot <- res_RFMix_plot_long |>
    filter(chr == chrnum)|> 
    ggplot(aes(x = pos, y = -log(Score.pval, 10), col = "RFMix")) +
    geom_point(data = res, aes(x = snppos, y =  -log(Score.pval, 10), 
                               col = "FLARE3"), 
               size = 0.6, alpha = 0.3) +
    geom_line(linewidth = 0.3) +
    labs(x = "") +
    scale_color_manual(values = colors,
                     name = legend_name) +
    theme_minimal()+
    theme(plot.title = element_text(hjust = 0.5),
          legend.position = "top") +
    guides(color = guide_legend(override.aes = list(size = 3, alpha = 0.8))) 
  return(overlay_plot)
}
```

**Use this function below to combine plots from different ancestries (African and Amerindian) and the two batches**
```{r}
# Combine batch 1 and batch 2 for both ancestries in an Miami plot
combine_plots  <- function(gafr_b1, gafr_b2, gamer_b1, gamer_b2,  title){
  gafr_b1 <- gafr_b1 + 
    theme(plot.margin = unit(c(0, 0.2, -2, 0.2), "lines"))
  
  gafr_b2 <- gafr_b2 + 
    theme(axis.text.x =element_blank(),
          plot.margin = unit(c(1, 0.2, 0, 0.2), "lines")) +
    theme(legend.position = "none")+
    scale_y_reverse()
  
  gamer_b1 <- gamer_b1 + 
    theme(plot.margin = unit(c(0, 0.2, -2, 0.2), "lines"))
  
  gamer_b2 <- gamer_b2 + 
    theme(axis.text.x =element_blank(),
          plot.margin = unit(c(1, 0.2, 0, 0.2), "lines")) +
    theme(legend.position = "none")+
    scale_y_reverse()
  
  # append the plots all together 
  grid.newpage()
  pushViewport(viewport(layout = grid.layout(2, 1, heights = unit(c(1.5, 8.5), "null"))))
  
  # Add the title
  grid.text(title, y = unit(1, "npc") - unit(1.5, "lines"), 
            gp = gpar(fontsize = 12, fontface = "bold"))
  
  # Add the combined plot
  pushViewport(viewport(layout.pos.row = 2, layout.pos.col = 1))
  grid.draw(cbind(rbind(ggplotGrob(gafr_b1),  ggplotGrob(gafr_b2)), 
                  rbind(ggplotGrob(gamer_b1),  ggplotGrob(gamer_b2)), 
                  size = "last"))
}
```

*Panel (A)*
Generate plots for 3-Aminoisobutyrate at chromosome 5, for both batches in FLARE3
```{r}
combine_plots(plot_overlay_pval("x1114", 1, "afr", 5), 
              plot_overlay_pval("x1114", 2, "afr", 5),
              plot_overlay_pval("x1114", 1, "amer", 5),
              plot_overlay_pval("x1114", 2, "amer", 5),
              "3-Aminoisobutyrate Chr5")
```

*Panel (B)*
Generate plots for N-acetylarginine at chromosome 2, for both batches in FLARE3
```{r}
combine_plots(plot_overlay_pval("x1266", 1, "afr", 2), 
              plot_overlay_pval("x1266", 2, "afr", 2),
              plot_overlay_pval("x1266", 1, "amer", 2),
              plot_overlay_pval("x1266", 2, "amer", 2),
              "N-acetylarginine Chr2")
```

*Panel (C)*
Generate plots for PE 16:0/20:4 at chromosome 15, for both batches in FLARE3
```{r}
combine_plots(plot_overlay_pval("x8990", 1, "afr", 15), 
              plot_overlay_pval("x8990", 2, "afr", 15),
              plot_overlay_pval("x8990", 1, "amer", 15),
              plot_overlay_pval("x8990", 2, "amer", 15),
              "PE 16:0/20:4 Chr15")
```

*Panel (D)*
Generate plots for PC 16:0/20:4 at chromosome 11, for both batches in FLARE3
```{r}
combine_plots(plot_overlay_pval("x8914", 1, "afr", 11), 
              plot_overlay_pval("x8914", 2, "afr", 11),
              plot_overlay_pval("x8914", 1, "amer", 11),
              plot_overlay_pval("x8914", 2, "amer", 11),
              "PC 16:0/20:4 Chr11")
```
