---
title: "Suppfig: Mean + Variance of ancestry proportions across HCHS/SOL individuals for each SNP (FLARE3) or ancestry interval (RFMix)"
author: "Xueying Chen"
date: "2024-07-15"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Import pacakges
```{r}
library(ggplot2)
library(RColorBrewer)
library(tidyverse)
```

# Load the correlation Rdata file
```{r}
#load("R:\\Sofer Lab\\HCHS_SOL\\Projects/2023_local_ancestry_comparison_sol/Data/FLARE3_RFMix_corr_by_chr/correlation_old_flare3_all_SNPs.Rdata") # a dataframe named dat_tot
load("/Volumes/Sofer Lab/HCHS_SOL/Projects/2023_local_ancestry_comparison_sol/Data/FLARE3_RFMix_corr_by_chr/correlation_old_flare3_all_SNPs.Rdata")

# load the combined files for SNPs post-QC (R2 > 0.8, MAF >= 0.005)
# filtered_snps <- readRDS("R:\\Sofer Lab\\HCHS_SOL\\Projects/2023_local_ancestry_comparison_sol/Data/FLARE3/FLARE3_SNPs_filtered_combined.rds")
filtered_snps <- readRDS("/Volumes/Sofer Lab/HCHS_SOL/Projects/2023_local_ancestry_comparison_sol/Data/FLARE3/FLARE3_SNPs_filtered_combined.rds")

# filter the SNPs for the correlation file
dat_tot <-  dat_tot |> inner_join(filtered_snps, by =  c("Chromosome" = "chr",
                                                         "Position" = "snppos"))

# total = 3339277 SNPs
nrow(dat_tot)/nrow(filtered_snps) # proportion of SNPs mapped to RFMix interval
```

# Calculating summary stats for mean ancestry proportions
```{r}
mean(dat_tot$Corr_afr_p)
sd(dat_tot$Corr_afr_p)
mean(dat_tot$Corr_amer_p)
sd(dat_tot$Corr_amer_p)
mean(dat_tot$Corr_eur_p)
sd(dat_tot$Corr_eur_p)
```

# Plot mean ancestry proportions, African ancestry, overlaying RFMix and FLARE3
```{r}
# Get color for plotting 
brewer.pal(n=3,"Dark2")

# Plotting global proportions across chromosomes (mean of all individuals), overlaying old and new
prop_afr <- dat_tot |> dplyr::select(Chromosome, Position, Prop_afr_old, Prop_afr_new)

# RFMix proportions - in intervals
prop_afr_old <- prop_afr |> group_by(Chromosome, Prop_afr_old) |>
  summarise(minpos = min(Position), maxpos = max(Position), .groups = "drop") 

prop_afr_old <- pivot_longer(prop_afr_old, cols = c("minpos", "maxpos"), 
                             names_to = "postype",
                            values_to = "pos") 

prop_afr_old <- prop_afr_old |> 
  arrange(Chromosome, pos)

g_prop_afr <- prop_afr |> ggplot(aes(x = Position, y = Prop_afr_new, 
                                     color = "FLARE3")) +
  geom_point(size = 0.3, alpha = 0.8) +
  geom_line(data = prop_afr_old,
            aes(x = pos, y = Prop_afr_old, color = "RFMix"), alpha = 0.8,
            linewidth = 0.5) +
  scale_color_manual(values = c("FLARE3" = "#AADCA9", "RFMix" = "#1B9E77"),
                     name = "Inference Version") +
  theme_minimal() +
  labs(y = "Mean Proportion - African Ancestry")+
  theme(axis.text.x =  element_blank(),
        axis.title.x = element_blank(),
        legend.position = "top",
        axis.title.y = element_text(size = 12),        
        axis.text.y = element_text(size = 10),         
        legend.title = element_text(size = 12),        
        legend.text = element_text(size = 10)) +
  guides(color = guide_legend(override.aes = list(size = 3, alpha = 1))) +
  facet_wrap(~Chromosome, scale = "free_x", ncol = 11) 

g_prop_afr
```

# Plot mean ancestry proportions, Amerinidan ancestry, overlaying RFMix and FLARE3
```{r}

prop_amer <- dat_tot |> dplyr::select(Chromosome, Position, Prop_amer_old, 
                                      Prop_amer_new)

# RFMix proportions - in intervals
prop_amer_old <- prop_amer |> group_by(Chromosome, Prop_amer_old) |>
  summarise(minpos = min(Position), maxpos = max(Position), .groups = "drop") 


prop_amer_old <- pivot_longer(prop_amer_old, cols = c("minpos", "maxpos"), 
                             names_to = "postype",
                            values_to = "pos") 

g_prop_amer <- prop_amer |> ggplot(aes(x = Position, y = Prop_amer_new,
                                       col = "FLARE3")) +
  geom_point(size = 0.2, alpha = 0.8) +
  labs(y = "Mean Proportion - Amerindian Ancestry")+
  geom_line(data = prop_amer_old, 
             aes(x = pos, y = Prop_amer_old, color = "RFMix"),
             linewidth = 0.5, alpha = 0.8) +
  theme_minimal() +
  scale_color_manual(values = c("FLARE3" = "#ffb38a", "RFMix" = "#D95F02"),
                     name = "Inference Version") +
  theme(axis.text.x =  element_blank(), 
        axis.title.x = element_blank(),
        axis.ticks.x = element_blank(),
        legend.position = "top",
        axis.title.y = element_text(size = 12),        
        axis.text.y = element_text(size = 10),         
        legend.title = element_text(size = 12),        
        legend.text = element_text(size = 10)) +
  guides(color = guide_legend(override.aes = list(size = 3, alpha = 1))) +
  facet_wrap(~Chromosome, scale = "free_x", ncol = 11)
g_prop_amer
# ggsave(g_prop_amer, file = "./prop_amer.pdf", width = 6, height = 4.5,
#        device = cairo_pdf)
```

# Plot mean ancestry proportions, European ancestry, overlaying RFMix and FLARE3
```{r}
prop_eur <- dat_tot |> dplyr::select(Chromosome, Position, Prop_eur_old, 
                                      Prop_eur_new)

# RFMix proportions - in intervals
prop_eur_old <- prop_eur |> group_by(Chromosome, Prop_eur_old) |>
  summarise(minpos = min(Position), maxpos = max(Position), .groups = "drop") 


prop_eur_old <- pivot_longer(prop_eur_old, cols = c("minpos", "maxpos"), 
                             names_to = "postype",
                            values_to = "pos")


g_prop_eur <- prop_eur |> ggplot(aes(x = Position, y = Prop_eur_new,
                                       col = "FLARE3")) +
  geom_point(size = 0.3, alpha = 0.8) +
  labs(y = "Mean Proportion - European Ancestry")+
  geom_line(data = prop_eur_old, 
             aes(x = pos, y = Prop_eur_old, color = "RFMix"),
             linewidth = 0.5, alpha = 0.8) +
  theme_minimal() +
  scale_color_manual(values = c("FLARE3" = "#CBC3E3", "RFMix" = "#7570B3"),
                     name = "Inference Version") +
  theme(axis.text.x =  element_blank(), 
        axis.title.x = element_blank(),
        axis.ticks.x = element_blank(),
        legend.position = "top",
        axis.title.y = element_text(size = 12),        
        axis.text.y = element_text(size = 10),         
        legend.title = element_text(size = 12),        
        legend.text = element_text(size = 10)) +
  guides(color = guide_legend(override.aes = list(size = 3, alpha = 1))) +
  facet_wrap(~Chromosome, scale = "free_x", ncol = 11)

# ggsave(g_prop_eur, file = "./prop_eur.pdf", width = 6, height = 4.5,
#        device = cairo_pdf)
g_prop_eur
```

# Plotting ancestry proportion variances across each SNP position

# African ancestry
```{r}
var_afr <- dat_tot |> dplyr::select(Chromosome, Position, Var_afr_old, Var_afr_new)

# RFMix ancestry proportion variances - in intervals
var_afr_old <- var_afr |> group_by(Chromosome, Var_afr_old) |>
  summarise(minpos = min(Position), maxpos = max(Position), .groups = "drop") 

var_afr_old <- pivot_longer(var_afr_old, cols = c("minpos", "maxpos"), 
                             names_to = "postype",
                            values_to = "pos")

g_var_afr <- var_afr |> ggplot(aes(x = Position, y = Var_afr_new,
                                       col = "FLARE3")) +
  geom_point(size = 0.3, alpha = 0.8) +
  labs(y = "Proportion  Variance - African Ancestry")+
  geom_line(data = var_afr_old, 
             aes(x = pos, y = Var_afr_old, color = "RFMix"),
             linewidth = 0.5, alpha = 0.8) +
  theme_minimal() +
  scale_color_manual(values = c("FLARE3" = "#AADCA9", "RFMix" = "#1B9E77"),
                     name = "Inference Version") +
  theme(axis.text.x =  element_blank(), 
        axis.title.x = element_blank(),
        axis.ticks.x = element_blank(),
        legend.position = "top",
        axis.title.y = element_text(size = 12),        
        axis.text.y = element_text(size = 10),         
        legend.title = element_text(size = 12),        
        legend.text = element_text(size = 10)) +
  guides(color = guide_legend(override.aes = list(size = 3, alpha = 1))) +
  facet_wrap(~Chromosome, scale = "free_x", ncol = 11)

g_var_afr

```

# Amerindian Ancestry
```{r}
var_amer <- dat_tot |> dplyr::select(Chromosome, Position, Var_amer_old, 
                                      Var_amer_new)

# RFMix ancestry proportion variances - in intervals
var_amer_old <- var_amer |> group_by(Chromosome, Var_amer_old) |>
  summarise(minpos = min(Position), maxpos = max(Position), .groups = "drop") 

var_amer_old <- pivot_longer(var_amer_old, cols = c("minpos", "maxpos"), 
                             names_to = "postype",
                            values_to = "pos")

g_var_amer <- var_amer |> ggplot(aes(x = Position, y = Var_amer_new,
                                       col = "FLARE3")) +
  geom_point(size = 0.3, alpha = 0.8) +
  labs(y = "Proportion  Variance - Amerindian Ancestry")+
  geom_line(data = var_amer_old, 
             aes(x = pos, y = Var_amer_old, color = "RFMix"),
             linewidth = 0.5, alpha = 0.8) +
  theme_minimal() +
  scale_color_manual(values = c("FLARE3" = "#ffb38a", "RFMix" = "#D95F02"),
                     name = "Inference Version") +
  theme(axis.text.x =  element_blank(), 
        axis.title.x = element_blank(),
        axis.ticks.x = element_blank(),
        legend.position = "top",
        axis.title.y = element_text(size = 12),        
        axis.text.y = element_text(size = 10),         
        legend.title = element_text(size = 12),        
        legend.text = element_text(size = 10)) +
  guides(color = guide_legend(override.aes = list(size = 3, alpha = 1))) +
  facet_wrap(~Chromosome, scale = "free_x", ncol = 11)

g_var_amer
```

#  European ancestry
```{r}
var_eur <- dat_tot |> dplyr::select(Chromosome, Position, Var_eur_old, 
                                      Var_eur_new)

# RFMix ancestry proportion variances - in intervals
var_eur_old <- var_eur |> group_by(Chromosome, Var_eur_old) |>
  summarise(minpos = min(Position), maxpos = max(Position), .groups = "drop") 

var_eur_old <- pivot_longer(var_eur_old, cols = c("minpos", "maxpos"), 
                             names_to = "postype",
                            values_to = "pos")

g_var_eur <- var_eur |> ggplot(aes(x = Position, y = Var_eur_new,
                                       col = "FLARE3")) +
  geom_point(size = 0.3, alpha = 0.8) +
  labs(y = "Proportion  Variance - European Ancestry")+
  geom_line(data = var_eur_old, 
             aes(x = pos, y = Var_eur_old, color = "RFMix"),
             linewidth = 0.5, alpha = 0.8) +
  theme_minimal() +
  scale_color_manual(values = c("FLARE3" = "#CBC3E3", "RFMix" = "#7570B3"),
                     name = "Inference Version") +
  theme(axis.text.x =  element_blank(), 
        axis.title.x = element_blank(),
        axis.ticks.x = element_blank(),
        legend.position = "top",
        axis.title.y = element_text(size = 12),        
        axis.text.y = element_text(size = 10),         
        legend.title = element_text(size = 12),        
        legend.text = element_text(size = 10)) +
  guides(color = guide_legend(override.aes = list(size = 3, alpha = 1))) +
  facet_wrap(~Chromosome, scale = "free_x", ncol = 11)

g_var_eur
```
